---
title: "SurvMap-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SurvMap-vignette}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette represents and introduction to the use of the package SurvMap. $\frac{m}{n}$ First we will show some aspects linked to matrix the denoising function implemented in the package, that is derived from (10.1109/TIT.2014.2323359). It allows to remove the noise part of matrices that are assumed to be composed by the addition of a signal matrix and a gaussian noise matrix.

```{r setup}
library(SurvMap)
```

Let's first create a signal matrix for testing.

```{r,fig.dim = c(5,7),fig.align = 'center'}
t <- seq(-3,3,0.01)
Utrue <- cbind(cos(17*t) * exp(-t^2), sin(11*t))
Strue <- cbind(c(2,0),c(0,0.5))
Vtrue <- cbind(sin(5*t)* exp(-t^2),cos(13*t))
X <- Utrue %*% Strue %*% t(Vtrue)
image(t(X))
```

Then add some gaussian noise to our matrix with $\sigma = 1$ and $\mu = 0$.

```{r,fig.dim = c(5,7),fig.align = 'center'}
sigma <- 1
noise <- matrix(rnorm(601 * 601),ncol = 601)
Xnoisy <- X + noise
image(t(Xnoisy))
```

Since we will work with rectangular matrices in most applications let's get a rectangular noisy matrix by removing some columns of our original noisy matrix.


```{r,fig.dim = c(5,7),fig.align = 'center'}
Xnoisy_r <- Xnoisy[,1:400]
dim(Xnoisy_r)
image(t(Xnoisy_r))
```

And lets do the same with the  signal matrix.

```{r,fig.dim = c(5,7),fig.align = 'center'}
X_r <- X[,1:400]
dim(X_r)
image(t(X_r))
```

Now let's  denoise our rectangular noisy matrix. 

```{r,fig.dim = c(5,7),fig.align = 'center'}
omega_found <- get_omega(ncol(Xnoisy_r)/nrow(Xnoisy_r))
svd_Xnoisy_r <- svd(Xnoisy_r)
D <- diag(svd_Xnoisy_r$d)
U <- svd_Xnoisy_r$u
V <- svd_Xnoisy_r$v
Xnoisy_reconstructed <- U %*% D %*% t(V)
image(t(Xnoisy_reconstructed))
```

```{r,fig.dim = c(5,7),fig.align = 'center'}
omega_found <- get_omega(ncol(Xnoisy_r)/nrow(Xnoisy_r))
svd_Xnoisy_r <- svd(Xnoisy_r)
D <- diag(svd_Xnoisy_r$d)
U <- svd_Xnoisy_r$u
V <- svd_Xnoisy_r$v
threshold_singular <- median(svd_Xnoisy_r$d)*omega_found
up_to_sv <- length(svd_Xnoisy_r$d[svd_Xnoisy_r$d > threshold_singular])
D_filt <- D
diag(D_filt)[(up_to_sv + 1):length(diag(D_filt))] <- 0
Xnoisy_denoised <- U %*% D_filt %*% t(V)
dim(Xnoisy_denoised)
image(t(Xnoisy_denoised))
```

```{r,fig.dim = c(5,7),fig.align = 'center'}
denoised_matrix <- denoise_rectangular_matrix(Xnoisy_r)


image(t(Xnoisy_r))
image(t(denoised_matrix))
is.matrix("t")
```

The data included in this Vignette is derived from GSE42568. A dataset that contains  
one hundred and twenty one samples. One hundred and four derived from breast cancer samples
and seventeen derived from healthy breast mammary tissues. 

```{r,eval=FALSE}
#Loading data.
data("GSE42568_Collapsed")
bool_filt <- pData_FRMA$pCh_Status == "NT"
Exp_NT <- GEO_Eset_Norm_frma_Collapsed[,bool_filt]
pData_NT <- pData_FRMA[bool_filt,]
Exp_T <- GEO_Eset_Norm_frma_Collapsed[,!bool_filt]
pData_T <- pData_FRMA[!bool_filt,]

#Generaling the healthy tissue model.

Exp_NT_f <- flatten_normal_tiss(Exp_NT)
class(Exp_NT_f)
Exp_NT_f_d <- denoise_rectangular_matrix(Exp_NT_f)
class(Exp_NT_f)


#cox_all <- cox_all_genes(Exp_T,pData_T$pCh_DFS_T,pData_T$pCh_DFS_E)


#Creating the disease component matrix.

#Disease_component <- generate_disease_component(GEO_Eset_Norm_frma_Collapsed,Exp_NT_f)
#Disease_component_tumors <- Disease_component[,!bool_filt]

#Selecting genes.

#variable_genes <- gene_selection(Disease_component_tumors,0.99) #Top variable genes in tumor samples.

#survival_related_genes <- get_survival_related_genes(cox_all,c(0.005,0.995)) #Genes showing the stringest association with disease-free survival in univariate proporciohal hazar models analysis. (cox.)

#selected_genes <- unique(c(survival_related_genes,variable_genes))
```

